---
title: "Reproducible Research - CP1"
author: "Taiguara Melo Tupinambás"
date: "7 de julho de 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r libraries}
library(plyr)
```

## Reading file

```{r reading data}
data<-read.csv("activity.csv",sep = ",", header = TRUE)
summary(data)
```

## What is mean total number of steps taken per day?

```{r mean total}
data$date<-as.Date(data$date)
actsum<-ddply(data, "date", summarize, 
              steps=sum(steps),
              interval=sum(interval))

hist(actsum$steps,breaks=10)

mean_steps<-mean(actsum$steps)
median_steps<-median(actsum$steps)

cat("Mean:",mean_steps,
    "\nMedian:",median_steps)
        
```

## What is the average daily activity pattern?

```{r avg patter}
dailypatt<-aggregate(steps~interval,
                     data=data,
                     FUN=mean)

plot(dailypatt,type="l",
     xlab="5-minute intervals",
     ylab="Average number of steps")

```

## Imputing missing values

### Counting missing values

```{r missing values}
NAs<-is.na(data$steps)
cat("Total number of missing values:",sum(NAs))
```

### Filling missing data

```{r filling data}

data_filled<-data

for (i in 1:length(data$steps)) {
        if (is.na(data[i,1])) {
                data_filled[i,1]<-dailypatt[which(dailypatt$interval==data[i,3]),2]
        }
}

NAs<-is.na(data_filled$steps)
cat("Total number of missing values:",sum(NAs))

```

### New data histogram, mean and median

```{r new data}


actsum_filled<-ddply(data_filled, "date", summarize, 
              steps=sum(steps),
              interval=sum(interval))

actsum_filled<-actsum_filled[complete.cases(actsum_filled),]

hist(actsum_filled$steps,breaks=10)

newmean_steps<-mean(actsum_filled$steps)
newmedian_steps<-median(actsum_filled$steps)

cat("New mean:",newmean_steps,
    "\nOld mean:",mean_steps,
    "\n\n\nNew Median",newmedian_steps,
    "\nOld Median:",median_steps)

```


By filling the missing data with the average number of steps for that 5-minute interval, the mean will not change, but the median will tend to get closer to the mean. The variability of the data is also expected to reduce.


## Are there differences in activity patterns between weekdays and weekends?


```{r weekday factor}

for (i in 1:length(data_filled$steps)) {
        if (weekdays(data_filled[i,2])=="sábado") {
                data_filled$day[i]<-"weekend"
        } else if (weekdays(data_filled[i,2])=="domingo") {
                data_filled$day[i]<-"weekend" 
        } else {
                data_filled$day[i]<-"weekday"
        }
}

```

```{r plot}

weekday<-data_filled[which(data_filled$day=="weekday"),]
weekend<-data_filled[which(data_filled$day=="weekend"),]

weekdaypatt<-aggregate(steps~interval,
                       data=weekday,
                       FUN=mean)
weekendpatt<-aggregate(steps~interval,
                       data=weekend,
                       FUN=mean)

par(mfrow=c(2,1))

plot(weekdaypatt,type="l",
     xlab="5-minute intervals",
     ylab="Avg # of steps Weekday")

plot(weekendpatt,type="l",
     xlab="5-minute intervals",
     ylab="Avg # of steps Weekend")

```

